---
title: "Team Attribute"
output:
  pdf_document: default
  html_document: default
date: "2026-10-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Data


```{r}
team_attributes <- read.csv("Team_Attributes.csv")
dim(team_attributes)
```
## Interaction Plot

```{r}
library(ggplot2)

ggplot(team_attributes,
       aes(x = chanceCreationPassing,
           y = chanceCreationShooting,
           color = chanceCreationPositioningClass)) +
  geom_point(alpha = 0.35) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    x = "Chance Creation Passing",
    y = "Chance Creation Shooting",
    color = "Positioning Class",
    title = "Passing vs. Shooting by Positioning Style"
  ) +
  theme_minimal()
```

## Data Preparation


```{r}
soccer_model_data <- team_attributes[, c(
  "chanceCreationShooting",
  "buildUpPlaySpeed",
  "buildUpPlayDribblingClass",
  "buildUpPlayPassing",
  "buildUpPlayPositioningClass",
  "chanceCreationPassing",
  "chanceCreationCrossing",
  "chanceCreationPositioningClass",
  "defencePressure",
  "defenceAggression",
  "defenceTeamWidth",
  "defenceDefenderLineClass"
)]

# Convert categorical variables to factors
soccer_model_data$buildUpPlayDribblingClass <-
  factor(soccer_model_data$buildUpPlayDribblingClass)

soccer_model_data$buildUpPlayPositioningClass <-
  factor(soccer_model_data$buildUpPlayPositioningClass)

soccer_model_data$chanceCreationPositioningClass <-
  factor(soccer_model_data$chanceCreationPositioningClass)

soccer_model_data$defenceDefenderLineClass <-
  factor(soccer_model_data$defenceDefenderLineClass)

# Check model dataset
dim(soccer_model_data)
colSums(is.na(soccer_model_data))
```
## Multiple Linear Regression Model

```{r}
soccer_model <- lm(
  chanceCreationShooting ~
    buildUpPlaySpeed +
    buildUpPlayDribblingClass +
    buildUpPlayPassing +
    buildUpPlayPositioningClass +
    chanceCreationPassing *
      chanceCreationPositioningClass +
    chanceCreationCrossing +
    defencePressure +
    defenceAggression +
    defenceTeamWidth +
    defenceDefenderLineClass,
  data = soccer_model_data
)

summary(soccer_model)
```
## Key Model Results

```{r}
# Full coefficient table
coef(summary(soccer_model))

# Model fit statistics
summary(soccer_model)$r.squared
summary(soccer_model)$adj.r.squared

# Sample size used in the model
nobs(soccer_model)
```
## Model Diagnostics

To assess whether the multiple linear regression assumptions are reasonably satisfied, diagnostic plots were examined for linearity, normality of residuals, constant variance, and potentially influential observations.

```{r}
# Model diagnostic plots

# 1. Residuals vs Fitted
plot(
  soccer_model,
  which = 1,
  main = "Residuals vs Fitted"
)

# 2. Normal Q-Q
plot(
  soccer_model,
  which = 2,
  main = "Normal Q-Q"
)

# 3. Scale-Location
plot(
  soccer_model,
  which = 3,
  main = "Scale-Location"
)

# 4. Residuals vs Leverage
plot(
  soccer_model,
  which = 5,
  main = "Residuals vs Leverage"
)
```
## Influential Observation Analysis

Potentially influential observations were evaluated using Cook's distance and leverage measures.

```{r}
# Calculate Cook's distance
cooks_d <- cooks.distance(soccer_model)

# Number of observations
n <- nobs(soccer_model)

# Common Cook's distance threshold
cook_threshold <- 4 / n

# Identify observations exceeding the threshold
influential_obs <- which(cooks_d > cook_threshold)

# Report threshold and number of potentially influential observations
cook_threshold
length(influential_obs)

# Display the 10 observations with the largest Cook's distances
head(sort(cooks_d, decreasing = TRUE), 10)

# Plot Cook's distance
plot(
  cooks_d,
  type = "h",
  main = "Cook's Distance",
  xlab = "Observation",
  ylab = "Cook's Distance"
)

abline(
  h = cook_threshold,
  lty = 2
)
```
## Interaction Effect Analysis

The interaction between chance-creation passing and positioning style was examined to determine whether the association between passing and shooting differs between Free Form and Organised teams.

```{r}
# Extract model coefficients
b <- coef(soccer_model)

# Estimated passing slope for Free Form teams
slope_freeform <- b["chanceCreationPassing"]

# Estimated passing slope for Organised teams
slope_organised <-
  b["chanceCreationPassing"] +
  b["chanceCreationPassing:chanceCreationPositioningClassOrganised"]

# Difference in passing slopes between positioning styles
slope_difference <-
  b["chanceCreationPassing:chanceCreationPositioningClassOrganised"]

# Display estimated slopes
slope_freeform
slope_organised
slope_difference

# 95% confidence intervals for the Free Form slope
# and the interaction coefficient
confint(soccer_model)[
  c(
    "chanceCreationPassing",
    "chanceCreationPassing:chanceCreationPositioningClassOrganised"
  ),
]
```
```{r}
# Statistical inference for the passing slope among Organised teams

V <- vcov(soccer_model)

b_pass <- coef(soccer_model)["chanceCreationPassing"]
b_int <- coef(soccer_model)[
  "chanceCreationPassing:chanceCreationPositioningClassOrganised"
]

# Estimated passing slope for Organised teams
slope_organised <- b_pass + b_int

# Standard error of the combined slope
se_organised <- sqrt(
  V["chanceCreationPassing", "chanceCreationPassing"] +
  V["chanceCreationPassing:chanceCreationPositioningClassOrganised",
    "chanceCreationPassing:chanceCreationPositioningClassOrganised"] +
  2 * V["chanceCreationPassing",
        "chanceCreationPassing:chanceCreationPositioningClassOrganised"]
)

# t statistic
t_organised <- slope_organised / se_organised

# p-value
p_organised <- 2 * pt(
  abs(t_organised),
  df = df.residual(soccer_model),
  lower.tail = FALSE
)

# 95% confidence interval
ci_organised <- slope_organised +
  c(-1, 1) *
  qt(0.975, df.residual(soccer_model)) *
  se_organised

# Display results
slope_organised
se_organised
t_organised
p_organised
ci_organised
```
```{r}
# Sensitivity analysis excluding observations flagged by Cook's Distance

# Create a sensitivity-analysis dataset
soccer_model_data_sensitivity <- soccer_model_data[-influential_obs, ]

# Refit the same model
soccer_model_sensitivity <- lm(
  chanceCreationShooting ~
    buildUpPlaySpeed +
    buildUpPlayDribblingClass +
    buildUpPlayPassing +
    buildUpPlayPositioningClass +
    chanceCreationPassing * chanceCreationPositioningClass +
    chanceCreationCrossing +
    defencePressure +
    defenceAggression +
    defenceTeamWidth +
    defenceDefenderLineClass,
  data = soccer_model_data_sensitivity
)

# Check remaining sample size
nobs(soccer_model_sensitivity)

# Display sensitivity-model results
summary(soccer_model_sensitivity)
```
```{r}
# Organised-team passing slope in the sensitivity model

V_sens <- vcov(soccer_model_sensitivity)

b_pass_sens <- coef(soccer_model_sensitivity)["chanceCreationPassing"]

b_int_sens <- coef(soccer_model_sensitivity)[
  "chanceCreationPassing:chanceCreationPositioningClassOrganised"
]

# Estimated passing slope for Organised teams
slope_organised_sens <- b_pass_sens + b_int_sens

# Standard error
se_organised_sens <- sqrt(
  V_sens["chanceCreationPassing", "chanceCreationPassing"] +
  V_sens[
    "chanceCreationPassing:chanceCreationPositioningClassOrganised",
    "chanceCreationPassing:chanceCreationPositioningClassOrganised"
  ] +
  2 * V_sens[
    "chanceCreationPassing",
    "chanceCreationPassing:chanceCreationPositioningClassOrganised"
  ]
)

# t statistic and p-value
t_organised_sens <- slope_organised_sens / se_organised_sens

p_organised_sens <- 2 * pt(
  abs(t_organised_sens),
  df = df.residual(soccer_model_sensitivity),
  lower.tail = FALSE
)

# 95% confidence interval
ci_organised_sens <- slope_organised_sens +
  c(-1, 1) *
  qt(0.975, df.residual(soccer_model_sensitivity)) *
  se_organised_sens

# Display results
slope_organised_sens
se_organised_sens
t_organised_sens
p_organised_sens
ci_organised_sens
```

